## For doing the rare and common variant association study, we first used the master VCF files called as Raw.Vcf 
## For this original file we divided the variaint files into 2 groups called as common and rare vcf files and did the manipulat##ion for the single variant assocaiation for the common variant and subsequenty planned to do the gene based test on the rare ##variant of the gene sets. 

## Part 1: Extracting the VCF files from the orignal raw files
First I made a new folder called Common_rare in the path 

/proj/b2012058/nobackup/private/sailen/

And copied the file raw.vcf into it

cp raw.vcf Common_rare/

## Now we the master VCF files contain samples that were unusable as from the mapping and subsequent preprocessing steps. We ## ##knew that there were some samples that had some problem with either the variants being low thorughput or with the samples are ##out of cluster from IBS plot (Contact Benjamin with the clustering plot). Thus, we only took the samples that had the high ##basecalls and grouped together so as to make their ethinic background same. For this we used the tools VCFtools. The samples ##in consideration were given a file name sample_remove and contains the sample names in the new files as
 
head sample_remove
S0328
S0664
S0580
S0922


## As the intial study was already done with the sample 211 with S1056 already excluded. While in the subsequent case-control study I took into the consideration the removal of sample S1056.

##For the common variants I selected the MAF 0.01 as the base line above which all the variants are selected while for the rare ## or in my opinion to take into the complete mutation profile such that even 1 allele has the variant, I took the MAF as 0.005 ## Now using the vcfools commands we get the following rare and common list variants from the raw vcf master files

 vcftools --vcf raw.vcf --remove sample_remove --remove-filtered-all --maf 0.01 --max-missing 0.95 --recode --out Common
 vcftools --vcf raw.vcf --remove sample_remove --remove-filtered-all --maf 0.005 --max-missing 0.90 --recode --out Rare
 ## THIS command did take a lot of time 
 
 ## Now we have the vcf file as Common.recode.vcf and Rare.recode.vcf and using the wc -l command in the each vcf file I get ##the following count of the vcf files

## Common = 74525 out of a possible 211467 Sites , 211 samples
## Rare =  91564 out of a possible 211467 Sites, 211 sample

## (See log files for the more information about the command and the output)

## Now having the two VCF files we proceed with the Single variant association and rare variant assocaition. For this I thought ## to use single variant association SNPs with the p value < 10-4 and mapping these variants into genes. These genes are then ## tested for the rare varaiant association. Does that make sense???

## Part -2 Associationg of the thrombopenia log nadir value with the variant using plinkseq linear regression model.
 
 module load plinkseq
 
 ## First I wanted to know the preliminary analysis of the individual variant calls: Hence I used the command
 
 pseq Common.recode.vcf i-stats | gcol ID NVAR SING TITV > Common_Individual_stat
 
 ## This showed me that the variant have the Ti/Tv ratio on average of 2.21 which I guess is not good but I will try to work ##with it and lets see if in the further downward process in case/Control there is improvement in the ratio.

##Making the proj files helps in the manipulation and managment of the vcf tools. Hence I created the project file into the ##plinkseq. But for creating the proj we needed the database files f the dbSNP, locus and refseq. So I copied the all the hg ##files and the data files from the previous folder.
 cp -r /Fom_Previous_File/hg19 /proj/b2012058/nobackup/private/sailen/Common_rare
 cp -r /From_Previos_File/data /proj/b2012058/nobackup/private/sailen/Common_rare
 
## However we needed the phenotype information and the phenotypic information was manipulated using R and the unix command  #lines  and stored in the data files such that
 
head data/TPK.phe 
 
##phe1,Float,-9,” Log modified Thrombopenia”
#ID	phe1
S0143	4.42
S0156	3.43
S0160	5.17
S0162	2.64
S0164	1.61
S0170	4.48
S0172	5.49
S0174	4.83
 
## Now making the project file in plinkseq

pseq proj new-project --resources hg19

## loading the file Common.recode.vcf in the plinkseq
pseq proj load-vcf --vcf Common.recode.vcf 
pseq proj tag-file --id 1 --name Common
pseq proj var-summary
pseq proj v-view

## Adding the phenotype information into the project
pseq proj load-pheno --file data/TPK.phe 
pseq proj i-view

## Doing the basic mandatory summary of the locus, database and seqbase

pseq proj loc-summary
pseq proj ref-summary 
pseq proj seq-summary

 pseq proj i-stats > Proj_istats
 
 ## Now after doing the preliminary analysis in plinkseq we wanted to do the linear regression of the quantitative value of the ## thrombopenia. For this we used the command glm in the plinkseq
 
 pseq proj glm --phenotype phe1 > Single_variant
 sort -g -k9 Single_variant > Sorted_Single_variant
 grep -v 'NA' Sorted_Single_variant > Final_Sorted_SNP
 wc -l Final_Sorted_SNP

## This is unadjusted linear regression of the quantitative endophenotype with the variants. We first did with the unajusted of ## the covariates as there may be compounding influence from the covariate. (See the paper "Including known covariates can  reduce #power to detect genetic effects in case-control studies")

#Now we extract the SNPs with the values <10-4

 sed -n "/VAR/,/0.000999356/p" Final_Sorted_SNP > Final_list_SNP
 wc -l Final_List_SNP  ## 89
 
 ## Now to map the Listed SNP to the genes or the near gene( because most of the genes are in the intronic region) I first uploaded the List into the Ensembl tool called as VEP. However here I noticed some discrepancy with the expected result. While slimming through the variant I had seen a gene SERPINC1 in the chromosome 1,which has seen function related to the thrombocytosis. However the ensembl shoed same position to be other gene so ask B about the tools or anything else that could map the variants to the REFSEQ because in the next phase of the association study we have database with the ref database. (See the miscellonous unix commands for the manipulation of the List to get to the final list.txt)

## In the mean time untill B is here I uploaded the file into the Ensembl database and extracted both the Ensembl VEP Gene Ids as well Ref sef Ids for the SNV from the single variant asssociation studies. Downloaded the files from the ensembl VEP and did the manipulation to get te both gene database. And finally stored the list online in the google doc at the moment. However it should be noted that the deletion SNV seen that could have effect are not mapped at the moment . Should talk about it as well. Additionally, there are couple of variant that don't mapped into refseq genes. 
And also there are genes ambigious symbols such as  AC015600.1, AL662800.2, C8orf37-AS1,CTD-2014B16.1, MIR1256, MIR324,RNU6-271P     RP11-172F4.2  RP11-253I19.3 while mapping using the Refseq symbols only. 
[57] RP11-342D11.2 RP11-382F24.1 RP11-513H8.1  RP11-599B13.9 RPS29P13

 However I found this tool called as SNPNexus that helped out great deal in getting the reference Gene for the refseq as well as the Ensemble. It is a online tool and required the files in the specific fomat. For which I did some manipulation in unix for the file. Then using the unique function in R I derived the gene list that mapped in these SNPs. It is named as the Gene_Refseq_TPK.
 
 ## Part -3 Rare-variant association for these specific genes:
 
 Doing some individual stats of the rare list
 For this we first annotated the rare variant list with the snpEff tool and using the command
 
 java -Xmx4g -jar /sw/apps/bioinfo/snpEff/4.1/milou/snpEff.jar -c /sw/apps/bioinfo/snpEff/4.1/milou/snpEff.config  -canon -hgvs -csvStats -s stats.test -d  GRCh37.75  /proj/b2012058/nobackup/private/sailen/Common_rare/Rare.recode.vcf  > /proj/b2012058/nobackup/private/sailen/Common_rare/Rare.Annotate.vcf
 
 Now in the rare list we have to get the vcf files of the genes that are seen in Refseq genes list we got from the previous analysis.To get the genes that are enriched with the gene variants with the genes we selected, I did the selection of the genes from the original Rare.Annoate.vcf file
 
 grep -f Gene_TPK.txt Rare.Annotate.vcf > Rare_to_be_test.vcf
 wc -l Rare_to_be_test.vcf 
 
 The original vcf file missed the header so we have to first extract the header and do the subsequent analysis so
 
 grep ^# Rare.Annotate.vcf > header.vcf
 cat header.vcf Rare_to_be_test.vcf > New.Rarelist.vcf
 
 
 ## 927 variant 
 
 
 ## The resst of the analysis for the SKAT was done in the local machine. For that I used the scp command and downloaded the file into the local machine. The following codes are run in the path
 
/Users/salendrapradh/Desktop 

where I have the plinkseq program 

## Creating the new project folder proj2 and doing he subsequent analysis in the project
plinkseq-0.10/pseq proj2 new-project --resources hg19
plinkseq-0.10/pseq proj2 load-vcf --vcf data/New.Rarelist.vcf 

plinkseq-0.10/pseq proj2 tag-file --id 1  --name Rare
plinkseq-0.10/pseq proj2 var-summary 

## Linking the project with the phenotype information of the TPK from the data folder
plinkseq-0.10/pseq proj2 load-pheno --file data/TPK.phe 
plinkseq-0.10/pseq proj2 i-view


## Doing the regular check ( sanity check) up of whether the databases are attached to the project or not

plinkseq-0.10/pseq proj2 loc-summary

## ---Locus DB summary---
## Group : refseq (41880 entries) n/a
## Group : ccds (27429 entries) n/a
## Group : gencode (94378 entries) n/a
## Group : ensembl (165825 entries) n/a

doing the similar stuffs for dbSNP and seqdB
plinkseq-0.10/pseq proj2 ref-summary 
##---Reference DB summary---
## Group : dbsnp (52054804 entries) : dbsnp (default name)

plinkseq-0.10/pseq proj2 seq-summary 

## Something like this should pop up
chr1:1..249250621	MB=249
chr2:1..243199373	MB=243
chr3:1..198022430	MB=198

## Deriving the summary statistics of the all the sample files
plinkseq-0.10/pseq proj2 v-stats > whole_project_summary.txt

## Trying the SKAT test for the refseq database 
 plinkseq-0.10/pseq proj2 assoc --tests skat --phenotype phe1 --mask loc.group=refseq  mac=1-10 > Rare_test_Variant_refseq.txt

## may be can be done for the ENSEMBl database as well
 plinkseq-0.10/pseq proj2 assoc --tests skat --phenotype phe1 --mask loc.group=ensembl  mac=1-10 > Rare_test_Variant_ensembl.txt

## now doing the manipulation with the derived files from the above 
 
## See the file Manipulation of SKAT_Quantitative to get the idea about the R 

## Finally we get the Files R_test_Ensembl and R_Test_Refseq to get the genes that are enriched. However note that these are not multiple test corrected. So may be that must be done.


